---
title: "Class Reference"
author: "Leonardo Micheli"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Consider this a personal guide to the commands and functions you will learn. In general, when you come across an R command or function that you want to remember, put it in here along with a description of what it does and when you'd use it.

### How to set the working directory

The command to set a working directory is setwd(). For example, this sets the working directory to a directory called data_journalism_2024_fall inside a directory called jour472 located inside your home directory (the ~ is equivalent of /Users/[youruser] on a Mac).

```{r}
setwd("~/jour472/data_journalism_2024_fall")
```

### How to install and load the tidyverse

```{r}
install.packages("tidyverse")
library(tidyverse)

### How to load a file from the Internet:
   # inside read_rds() function, we’ve put the name of the file we want to load

maryland_winred_contributions <- read_rds("https://thescoop.org/files/maryland_winred.rds")

### How to have a first view of the file

glimpse(maryland_winred_contributions)
or
head(maryland_winred_contributions)

# take your data set and then do this specific action to it.
  data |> function

### How to count data
  |> group_by(COLUMN NAME) |> 
    summarize(VARIABLE NAME = AGGREGATE
      FUNCTION(COLUMN NAME))
    e.g.:
maryland_winred_contributions |>
  group_by(memo_text) |>
  summarise(
    count_contribs = n()
  )

# How to count the number oof rows or records in each group
    n()

### How to arrange data in order
    arrange()
maryland_winred_contributions |>
  group_by(memo_text) |>
  summarise(
    count_contribs = n()
  ) |>
  arrange(desc(count_contribs))

### How to know the total amount of a set of data
  We put the column we want to total – “amount” – inside the sum() function sum(amount)
  e.g.:
maryland_winred_contributions |>
  group_by(memo_text) |>
  summarise(
    count_contribs = n(),
    total_amount = sum(amount)
  ) |>
  arrange(desc(total_amount))
  
### How to calculate the average amount for each set of data - the mean - and the amount that sits at the midpoint of our data - the median.
maryland_winred_contributions |>
  group_by(memo_text) |>
  summarise(
    count_contribs = n(),
    total_amount = sum(amount),
    mean_amount = mean(amount),
    median_amount = median(amount)
  ) |>
  arrange(desc(count_contribs))

### How to calculate the highest and lowest set of data
 we can use the min() and max() functions.
e.g.:
 maryland_winred_contributions |>
  group_by(memo_text) |>
  summarise(
    count_contribs = n(),
    total_amount = sum(amount),
    mean_amount = mean(amount),
    median_amount = median(amount),
    min_amount = min(amount),
    max_amount = max(amount)
  ) |>
  arrange(desc(max_amount)) 

assignment 08/30 https://github.com/lmicheli21/data_journalism_2024_fall/commit/26e97ce97ae7a8720346ec501cb3b0c6573fa82e