---
title: "lab_02"
author: "Leonardo Micheli"
date: "2024-08-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
library(lubridate)
```

## Load Data

For this lab we'll be working with a slice of Maryland voter registration data: records of people who have registration dates this year through Aug. 7. Let's load it.

```{r}
md_voters_2024 <- read_rds("data/lab_02.rds")
```

## Examine Data

Take a look at the data, getting familiar with what it contains. Most of the columns should be self-explanatory, but there are two date columns: STATE_REGISTRATION_DATE and COUNTY_REGISTRATION_DATE. The former represents the date the person registered as a Maryland voter; the latter is when they registered in their current county. They may be the same, but if a voter moves from one county to another they will differ. This is important because if the two dates are the same, that's a pretty good indication that this is a *new* voter, someone who has no history of registering in Maryland. You'll need to keep that definition in mind.

## Answer questions

**Q1** How many voters appear to be *new* registrants since July 21, when President Joe Biden announced he would leave the race, based on the information above? Create a dataframe to hold the only those voters. Which counties have seen the most? Take a look at county population figures here and compare the two in your answer: <https://worldpopulationreview.com/us-counties/maryland>.

**A1** The top 3 counties with the most new rigistrants are Montgomery (4847) Baltimore City (2799) Baltimore County (1683). In comparison to the county population, Prince George County doesn't result in the top 3 counties with most new registrants (4th for most new registrants, while 2nd for most inhabitants).

```{r}
filtered_md_voters_2024 <- md_voters_2024 |> filter(STATE_REGISTRATION_DATE == COUNTY_REGISTRATION_DATE)

head(filtered_md_voters_2024)
```

```{r}

Biden_filtered_md_voters_2024 <- filtered_md_voters_2024 |> filter(STATE_REGISTRATION_DATE > as_date("2024-07-21"))

head(Biden_filtered_md_voters_2024)

```

```{r}
Biden_filtered_md_voters_2024 |> 
  group_by(COUNTY) |>
  summarise(
  COUNTY_TOTAL  = n()) |>
  arrange(desc(COUNTY_TOTAL))
```



**Q2** A better question is whether the new voters you found in Q1 are an outlier compared to other weeks this year. Using the original dataframe with 2024 registrants, use `mutate()` and lubridate to produce week numbers from the COUNTY_REGISTRATION_DATE (Lubridate has a function that turns a date into a week number. Use that to answer this question). Save that to the original dataframe. Then calculate how many voters are in each week. Describe the distribution of voters groups in a sentence - are the weeks from July 21-August 7 different from the rest, and if so, how?

**A2** The voters between the 29th week and the 32nd were 20966, while for the rest they were 89470. This means that the registrants per week during the weeks when President Joe Biden announced he would leave the race were much more than for rest of the time's week. 

```{r}
md_voters_2024 |>
  mutate(week = week(STATE_REGISTRATION_DATE)) |>
  group_by(week) |>
  summarise(week_totals = n()) |>
  filter(week >= 29) |>
  summarise(
    votes_per_week_total_29_32 = sum(week_totals)
  ) 

md_voters_2024 |>
  mutate(week = week(STATE_REGISTRATION_DATE)) |>
  group_by(week) |>
  summarise(week_totals = n()) |>
 filter(week<29) |>
  summarise(
    votes_rest = sum(week_totals)   
  )

```

```{r}
week('2024-07-21')
week('2024-08-07')
```


**Q3** Using lubridate, show which month saw the largest number of registrations. Does the answer help shed more light on Q2? 
**A3** the month that saw the largest number of registrations was July (26139), following up by June (19475) and March (15960). Once again, we can notice how Presiden Biden's decision made people register more. 

```{r}
md_voters_per_month_ <- md_voters_2024 |>
  mutate(month = month(STATE_REGISTRATION_DATE)) |>
  group_by(month) |>
  summarise(voters_per_month = n()) |>
  arrange(desc(voters_per_month))

```



**Q4** Maryland has, officially, five regions: <https://www.visitmaryland.org/article/maryland-regions>. Modify the original dataframe again to add a `region` column based on the `COUNTY` value using case_when. Are you going to have to do this for every Maryland county plus Baltimore City? Yes, you are. Which regions and parties account for the most new voters?

**A4** The top 4 regions and parties that account for the most new voters are the DEM in the Capital Region, UNA in Capital Region, UNA in Central Maryland, and DEM in Central Maryland.  

```{r}

region_md_votest <- md_voters_2024 |>
  mutate(
    Region = case_when(
        COUNTY == "Allegany" | COUNTY == "Garrett" | COUNTY == "Washington" ~ "Western Maryland",
        COUNTY == "Baltimore County" | COUNTY == "Carroll" | COUNTY == "Harford" | COUNTY == "Howard" | COUNTY == "Anne Arundel" ~ "Central Maryland",
        COUNTY == "Prince George's" | COUNTY == "Montgomery" | COUNTY == "Frederick" ~ "Capital Region",
        COUNTY == "Calvert" | COUNTY == "Charles" | COUNTY == "St. Mary's" ~ "Southern Maryland",
        COUNTY == "Kent" | COUNTY == "Queen Anne's" | COUNTY == "Talbot" | COUNTY == "Caroline" | COUNTY == "Dorchester" | COUNTY == "Wicomico" | COUNTY == "Somerset" | COUNTY == "Worcester" | COUNTY == "Cecil" ~ "Eastern Shore",
        COUNTY == "Baltimore City" ~ "Baltimore City"
    )
  )


```

```{r}
filtered_region_md_votest_ <- filtered_md_voters_2024 |>
  mutate(
    Region = case_when(
        COUNTY == "Allegany" | COUNTY == "Garrett" | COUNTY == "Washington" ~ "Western Maryland",
        COUNTY == "Baltimore County" | COUNTY == "Carroll" | COUNTY == "Harford" | COUNTY == "Howard" | COUNTY == "Anne Arundel" ~ "Central Maryland",
        COUNTY == "Prince George's" | COUNTY == "Montgomery" | COUNTY == "Frederick" ~ "Capital Region",
        COUNTY == "Calvert" | COUNTY == "Charles" | COUNTY == "St. Mary's" ~ "Southern Maryland",
        COUNTY == "Kent" | COUNTY == "Queen Anne's" | COUNTY == "Talbot" | COUNTY == "Caroline" | COUNTY == "Dorchester" | COUNTY == "Wicomico" | COUNTY == "Somerset" | COUNTY == "Worcester" | COUNTY == "Cecil" ~ "Eastern Shore",
        COUNTY == "Baltimore City" ~ "Baltimore City"
    )
  )
```


```{r}
registrations_by_region_new_voters <- filtered_region_md_votest_ |>
  group_by(Region) |>
  summarise(number_of_registrations = n())

```

```{r}
registrations_by_region_and_party_new_voters <- filtered_region_md_votest_ |>
  group_by(Region, PARTY) |>
  summarise(number_of_registrations = n())
```

```{r}
registrations_by_region_and_party_new_voters <- filtered_region_md_votest_ |>
  group_by(Region, PARTY) |>
  summarise(number_of_registrations = n()) |>
arrange(desc(number_of_registrations))
```


**Q5** What's an additional question you could ask of this data?

**A5** Create a new column and put a category value based on the amount of each registrations per week

```{r}
largest_md_votest <- md_voters_per_week_ |>
  mutate(
    amount_registrations = case_when(
        voters_per_week < 1000 ~ "Less than 1000",
        voters_per_week >= 1000 & voters_per_week < 2000 ~ "Between 1000 and 1999",
        voters_per_week >= 2000 & voters_per_week < 3000 ~ "Between 2000 and 2999",
        voters_per_week >= 3000 & voters_per_week < 4000 ~ "Between 3000 and 3999",
        voters_per_week >= 4000 & voters_per_week < 5000 ~ "Between 4000 and 4999",
        voters_per_week >= 5000 & voters_per_week < 6000 ~ "Between 5000 and 5999",
        voters_per_week >= 7000 ~ "7,000 or more"
      )
  )
head(largest_md_votest)
```

