---
title: "lab_03"
author: "Leonardo Micheli"
date: "2024-08-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse
library(tidyverse)

# Load janitor
library(janitor)
```

## Load data

```{r}
# Load MD slice of FEC House candidate contribution data
md_contribs <- read_rds("data/lab_03.rds")
```

## Fix header issues

You'll need to fix one issue with the column names using janitor. Write a code block to do that, making sure that you save the results to the same dataframe.

```{r}
md_contribs <- md_contribs |>
  clean_names()
```


## Answer questions

Q1: Which seven cities in Maryland had the largest *number* of contributions from donors? How many did each have? To answer this question, you'll have to clean the messy "contributor_city" field.

Note: To answer this question, you'll need to see every variation of the "contributor_city" for the top seven cities, which you should identify through some initial analysis. Then find variations of them and change those variations into the most common version.


A1: The seven cities in Maryland that have the largest number of contributions from donors are BETHESDA	(1177) CHEVY CHASE	(704), BALTIMORE	(690), ANNAPOLIS (490), POTOMAC	(490), SILVER SPRING (405), ROCKVILLE	(358)	

```{r}
md_contribs |>
  group_by(contributor_city) |> 
  summarise(count = n()) |>
 arrange(desc(contributor_city))

```

```{r}
md_contribs |> 
  get_dupes()
```
# annapolis --> anapolis 
# Baltimore --> balrimore, blatimore
# Bethesda 
# Chevy Chase 
# Potomac --> potamic
# Rockville --> rockwell
# Silver spring --> silver springs

```{r}
cleaned_md_contribs <- md_contribs |>
  mutate(contributor_city = case_when(
    contributor_city == "ANAPOLIS" ~ "ANNAPOLIS",
    contributor_city == "BALRIMORE" ~ "BALTIMORE",
    contributor_city == "BLATIMORE" ~ "BALTIMORE",
    contributor_city == "POTAMIC" ~ "POTOMAC",
    contributor_city == "ROCKVILLE" ~ "ROCKWELL",
    contributor_city == "SILVER SPRINGS" ~ "SILVER SPRING",
        TRUE ~ contributor_city
  ))

```


```{r}
md_contribs |>
  group_by(contributor_city) |> 
  summarise(count = n()) |>
 arrange(desc(count))
```


*Q2*: Find the 5-digit ZIP code in Maryland that had the highest total amount of contributions (by dollar value). You will need to use mutate to ensure that you have a zip code column with the first five characters from the existing zip code column.

Where is that zip code and what kind of area is it, demographically? Which committee is the largest recipient of money from that zip code? How many of the top five recipients are Maryland candidates? 

A2:The ZIP code that has the highest total amount of contributions is 20817, localized in Bethesda. According to censusreporter.org, the per capita income is $78,300. Overall, it seems like a wealthier area. The committee with the largest recipient of money from that zip code is HELMER FOR CONGRESS. Within the top 5 recipients, the only one that are Maryland are "JAMIE RASKIN FOR CONGRESS" and "APRIL MCCLAIN DELANEY FOR CONGRESS". 

```{r}
zip_md_contribs <- cleaned_md_contribs |>
  mutate(contributor_zip = str_sub(contributor_zip, start=1L, end=5L))
```


```{r}
zip_md_contribs |>
  mutate(contributor_zip = str_sub(contributor_zip, start=1L, end=5L)) |>
  group_by(contributor_zip) |> 
  summarise(total_contributions = sum(contribution_receipt_amount)) |>
 arrange(desc(total_contributions))
```

```{r}
total_zip_md_contribs <- zip_md_contribs |>
  filter(contributor_zip == "20817") |>
  group_by(committee_name) |> 
  summarise(total_contributions = sum(contribution_receipt_amount)) |>
 arrange(desc(total_contributions))
```


Q3: Let's look at top individual donors. Write code to generate a list of top donors using name and city, adding up the total amount of contributions. What does the original data say about the top two donors - who are they and what do they do? Do you see anything in the list of top 10 donors that would make you uncertain about the accuracy of these results?

A3: The top two donors are Melanie Jacobsen and Daniel Kracov. Melanie, the largest contributor, has donated $96,466.65 and is listed as a farmer. Daniel, the second-largest contributor with $55,499.80, is an attorney. However, when examining the top 10 donors, it's notable that many are listed as "not employed" or "retired." This raises some concerns about data accuracy. For instance, Mark Joseph, one of the top contributors, shows inconsistencies in his records. Out of 17 contributions, his occupation is listed as "retired" in one instance and "not employed" in the others, which suggests potential data entry errors or inconsistencies.


```{r}
individual_donors <- cleaned_md_contribs |>
  group_by(contributor_name, contributor_city) |>
  summarise(total_contributions = sum(contribution_receipt_amount)) |>
  arrange(desc(total_contributions))

```

```{r}
who_individual <- cleaned_md_contribs |>
   group_by(contributor_name, contributor_city, contributor_occupation, ) |>
  summarise(total_contributions = sum(contribution_receipt_amount)) |>
  arrange(desc(total_contributions))
```

