---
title: "lab_01"
author: "Leonardo Micheli"
date: "2024-08-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Turn off scientific notation
options(scipen=999)
```

## About this lab

To complete this lab, you need to: \* run existing code as directed (look for **Task**). \* modify existing code as directed (look for **Task**). \* write code in empty codeblocks provided to answer questions included (look for **Q**). \* write out the answer in the form of a complete sentence in the space given (look for **A**). When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

Remember to add any new verbs or concepts to your reference notebook!

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
```

## Load Data

We're going to load the slice of WinRed contribution data from West Virginia, which is in a file called `lab_01.rds`

**Task** Using the example from the pre_lab, write code to load the RDS file and assign it to a good variable name.

```{r}
westvirginia_winred_contributions <- read_rds("lab_01.rds")
```



## Examine Data

**Task** Run the codeblock below to use glimpse to get a sense of the column names, types and sample data.ù
```{r}
glimpse(westvirginia_winred_contributions)
```

**Q1** How many rows and columns does this dataframe have?

**A1** Thanks to the glimpse function we can see that there are 40,441 rows and 24 columns

```{r}
westvirginia_winred_contributions |>
  group_by(city) |>
  summarise(
    count_contribs = n()
  ) |>
  arrange(desc(count_contribs))
```

## Answer questions

**Q2** Which West Virginia city provided the largest number of contributions? How many?

**A2** Charleston is the West Virginia city that provided the largest number of contributions with 3226 total contributions

```{r}
westvirginia_winred_contributions |>
  group_by(city) |>
  summarise(
    count_contribs = n()
  ) |>
  arrange(count_contribs)
```

**Q3** Which West Virginia cities accounted for the fewest contributions? How many of them, roughly speaking?

**A3** There are different cities that account for only 1 contributions such as Allen Junction, Alum Bridge and Bergoo.Overall, there are 73 cities that account for 1 contribution. 


**Q4** What's your best educated guess to explain why the city identified in question 2 had more contributions compared to the cities identified in question 3? Hint: you won't find the answer in the data, you'll need to do some quick web research. You do not need to write any code for this question, so no codeblock is provided.

**A4** Charleston, identified in Question 2, is the most populous city in West Virginia, which naturally accounts for the highest number of contributions. In contrast, the cities mentioned in Question 3 represent some of the state's least populated towns, often with fewer than 100 residents. For instance, Bergoo, according to Google, has only 93 inhabitants, making it understandable that it has only a single contribution. 

**Q5** What recipient got the largest total amount of money? How many contributions did they receive and what was the average donation?

**A5** The recipient got the largest total amount of money is "Earmarked for TRUMP NATIONAL COMMITTEE JFC, INC. (C00873893)" with 14,063 contributions. Totally they received 463944.40 dollars. And the average donation was 32.9904288 dollars. 	

```{r}
westvirginia_winred_contributions |>
  group_by(memo_text) |>
  summarise(
    count_contribs = n(),
    total_amount = sum(amount),
    mean_amount = mean(amount),
    median_amount = median(amount)
  ) |>
  arrange(desc(total_amount))
```

**Q6** What is that top recipient and what explains its popularity? Hint: this will require web research, you won't be able to answer this purely from the data set. You do not need to write any code for this question, so no codeblock is provided.

**A6** As reported by CNN, West Virginia has consistently voted Republican since the 2000s. Therefore, it makes sense that the most popular recipient in this case study aligns with the most popular Republican candidate: Donald Trump. 

**Q7**. Using the answer from A5 as a starting point, produce a summary table that has the following information for each recipient: \* count of contributions \* total amount of all contributions \* the average contribution amount \* the median contribution amount \* the smallest contribution amount \* the largest contribution amount

Store it as a variable object called `wv_summary` so that you can examine it with the data viewer. What newsworthy nuggets jump out at you that you could imagine dropping into a news story? Write a few summary sentences with overall findings.

**A7** 
The city with the highest average contribution amount, among those with more than 900 contributors, is Morgantown, with an average of $42.06, highlighting its strong financial support. Interestingly, the top three cities in total contributors - Wheeling, Martinsburg, and Morgantown  all come from the northern part of the state, with Charleston being the exception. This north-south divide could be worth exploring further to see if it influences other factors. Additionally, smaller towns like Pecks Mill, Bethlehem, Northspring, and Hometown stand out for having the highest average contribution amounts, indicating concentrated financial support in these areas. 


```{r}
wv_summary <- westvirginia_winred_contributions |>
  group_by(city) |>
  summarise(
    count_contribs = n(),
    total_amount = sum(amount),
    mean_amount = mean(amount),
    median_amount = median(amount),
    min_amount = min(amount),
    max_amount = max(amount)
  ) |>
  arrange(desc(total_amount))
```
