---
title: "Reverse Engineering Project"
author: "Student names here"
date: "Date here"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [STORY NAME](STORY URL)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
options(scipen=999) 
library(tidyverse)
library(tidyverse)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data
congress_age <- read_csv("data/data_aging_congress.csv")
census_2020_age <- read_csv("data/census_2020_age.csv")

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 
glimpse(congress_age)

```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: [According to the 2020 census, the median age of the entire U.S. population was about 39, and among those 25 and older, it was 51 — almost 22 and **10 years younger, respectively, than the median member of the 116th Congress was at the time.**]
* **Analysis summary**: [We were able to confirm that the 116th congress is almost 22 years older than the census data of the entire population, i.e., 39+22 = 61. Our data reveals the median age to be 59. We looked at the census data for 2020. src: https://data.census.gov/table/ACSST5Y2020.S0101?q=age%20in%202020 imported as census_2020_age.csv in 'data' folder. The census data explicitly mentions 39 to be the median data in their csv table.]

```{r}
# Put code to reverse engineer sentence here
congress_age |>
  filter(congress == "116") |>
  summarise(
    med_age_116 = median(age_years)
  )
# Display results of code below this codeblock
58.98563
```

### Sentence 2

* **Sentence text**: [According to the 2020 census, the median age of the entire U.S. population was about 39, and **among those 25 and older, it was 51 **]
* **Analysis summary**: [We imported the census 2020 data and wanted to check the median age for people 25 and above. The United States!!Total!!Estimate column has the population for each age group. However, these have commas in them and the code reveals its a in character format. We removed the commas/any non numeric data and coerced it into a numeric fata type. However, filtering just the 25 and above data groups returns a file with no observations. Not sure why that is. Maybe we would also have to account for how many people there are in each age group to find a weighted median, unsure how to proceed with that.]

```{r}
# Put code to reverse engineer sentence here
glimpse(census_2020_age) #the The United States!!Total!!Estimate column is chr 
census_2020_age <- census_2020_age %>%
  mutate(
    Population = as.numeric(str_remove_all(`United States!!Total!!Estimate`, "[^0-9]")) #removing anything that is not a digit
  )
str(census_2020_age) #population is now numeric without commas 
data_filtered <- census_2020_age %>%
  filter(`Label (Grouping)` %in% c(
    "25 to 29 years", "30 to 34 years", "35 to 39 years", 
    "40 to 44 years", "45 to 49 years", "50 to 54 years",
    "55 to 59 years", "60 to 64 years", "65 to 69 years", 
    "70 to 74 years", "75 to 79 years", "80 to 84 years", 
    "85 years and over"
  ))
# Display results of code below this codeblock
```

### Sentence 3

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 4

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-